---
title: 'Military Deployment visualizations for Article'
author: "Emil Staševski"
output: 
  html_document:
    theme: paper
    code_folding: hide
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading libraries


```{r, message=FALSE,warning=FALSE}
library(tidyverse)
library(tidyr)
library(readxl)
library(stargazer)
library(dplyr)
library(knitr)

colors3 <- c("#e6194b", "#3cb44b", "#ffe119", "#4363d8", "#f58231", "#911eb4", "#46f0f0", "#f032e6", "#bcf60c",
             "#fabebe", "#008080" ,"#e6beff" ,"#9a6324" ,"#fffac8", "#800000", "#aaffc3", "#808000", "#ffd8b1",
              "#000075" ,"gold"  ,  "#ffffff", "#000000")
```



### Loading & joining data

```{r, message=FALSE,warning=FALSE}

balance = read.csv(url("https://raw.githubusercontent.com/EmilStasevski/Military-Deployment/main/IMDT.csv"))


balance$year_ccode1 = paste(balance$year, balance$ccode1, sep='_')

corwar = read.csv(url("https://raw.githubusercontent.com/EmilStasevski/Military-Deployment/main/Correlates_of_war.csv"))
corwar$year_ccode1=paste(corwar$year, corwar$ccode1, sep='_')


corwar %>% select(military_personnel, year_ccode1)->corwar
balance %>% left_join(corwar)->balance

```


#### Creating features

```{r, message=FALSE,warning=FALSE}
balance$period <- case_when(
  balance$year<1993 ~ '1985-1992',
  balance$year>1992 & balance$year< 2001~ '1993-2000',
  balance$year>2000 & balance$year <2014~'2001-2013',
  balance$year>2013 ~ '2014-2022'
)

balance$troops =as.numeric(balance$Troops)

balance$percent=balance$troops / balance$military_personnel
```





### Figure 1

*The military deployment ratio to total military personnel in 1985-2022*


```{r, message=FALSE,warning=FALSE}
pathImages= '~/military deployment/images/'


balance %>% 
  group_by(year) %>% summarise(percent = mean(percent, na.rm=T)) %>% 
  ggplot(aes(year,percent ))+
  geom_line(size=1,linetype = "dashed",color='blue')+
  labs(title = 'Percent of military personnel in 1985-2022',y='Percent',x='Year')+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))+
  theme_bw()+
   theme(axis.title.x = element_text(size=8, face='bold'),axis.title.y = element_text(size=8, face='bold'), title =element_text(size=10, face='bold') )+
  theme(axis.text.x = element_text( size=6))+
  theme(axis.text.y = element_text( size=6))


ggsave('~/Articles/images/Figure1.jpg', dpi=800)
ggsave('~/Articles/images/Figure1.eps', dpi=800)
```


### Figure 2 

*Confidence interval for mean values of deployed military personnel down by periods*


```{r, message=FALSE,warning=FALSE}


df <- balance %>% group_by(period) %>% summarise(m = mean(troops, na.rm=T),
                                       stdv = sd(troops, na.rm=T))
df$period <- as.factor(factor(df$period, levels=c("1985-1992",     
 "1993-2000","2001-2013" ,"2014-2022" )) )


ggplot(df, aes(period,m,fill=period)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=m-stdv, ymax=m+stdv), width=.2,
                position=position_dodge(.9)) +
  labs(title = 'Confidence interval for means',y='Percent',x='Period')+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))+
  theme_bw()+
   theme(axis.title.x = element_text(size=8, face='bold'),axis.title.y = element_text(size=8, face='bold'), title =element_text(size=10, face='bold') )+
  theme(axis.text.x = element_text( size=6))+
  theme(axis.text.y = element_text( size=6))


ggsave('~/Articles/images/Figure2.jpg', dpi=800)
ggsave('~/Articles/images/Figure2.eps', dpi=800)
```



### Figure 3

*Top 10 biggest deployers by quantity of deployed troops and top 10 most frequent deployers in 1985-2022*


```{r, message=FALSE,warning=FALSE,fig.width=10,fig.height=4}
balance %>% 
  group_by(deployer) %>% summarise(troops=mean(troops, na.rm=T)) %>%  arrange(-troops) %>%  head(10) %>% select(deployer) %>% left_join(balance, by='deployer') %>% 
  group_by(deployer, year) %>% summarise(troops=mean(troops, na.rm=T)) %>% 
  ggplot(aes(year, troops, fill=deployer))+
  geom_bar(stat='identity')+
  labs(y='Troops',x='Year')+
  scale_fill_manual(values= colors3)+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))+
  theme_bw()+
   theme(axis.title.x = element_text(size=8, face='bold'),axis.title.y = element_text(size=8, face='bold'), plot.title =element_text(size=14, face='bold') )+
  theme(axis.text.x = element_text( size=6))+
  theme(axis.text.y = element_text( size=6))->p1

 


balance %>% 
  group_by(deployer) %>% summarise(n=n()) %>%  arrange(-n) %>%  head(10) %>% select(deployer) %>% left_join(balance, by='deployer') %>% 
  group_by(deployer, year) %>% summarise(n=n()) %>% 
  ggplot(aes(year,n, fill=deployer))+
  geom_bar(stat='identity')+
  labs(y='Number of dyads',x='Year')+
  scale_fill_manual(values= colors3)+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))+
  theme_bw()+
   theme(axis.title.x = element_text(size=8, face='bold'),axis.title.y = element_text(size=8, face='bold'), title =element_text(size=10, face='bold') )+
  theme(axis.text.x = element_text( size=6))+
  theme(axis.text.y = element_text( size=6))->p2


library(ggpubr)
figure <- ggarrange(p1,p2)
annotate_figure(figure, top = text_grob("Top deployers", 
               face = "bold", size = 14))




ggsave('~/Articles/images/Figure3.jpg', dpi=800, width=20, heigh=8)
ggsave('~/Articles/images/Figure3.eps', dpi=800, width=20, heigh=8)

```



### Figure 4


*Military deployment down by mission types in four periods*


```{r, message=FALSE,warning=FALSE}
balance %>% 
  select(ccode1,ccode2,troops,deployer,period,host
        
         ,c(UN:Air))->operations

operations %>% 
  pivot_longer(cols=c('Peacekeeping':'Air'),
                names_to='Type',
               values_to = 'freq')->longed_type
```



```{r, message=FALSE,warning=FALSE}

group.colors <- c(Peacekeeping = "#4363d8",  Other="gray", 
       no_mandate='#303030',           
  CounterInsurgency ='red',   
  Training='orange',
  Air='cyan3',
  Maritime='navyblue',
  no_type='gray'
  
 )

longed_type$Type = ifelse(longed_type$freq<1 , 'no_type', longed_type$Type)

longed_type %>% 
  group_by(period, Type) %>% summarise(troops=mean(troops,na.rm=T)) %>% 
  arrange(desc(period)) ->groupped

groupped$Type=ifelse(is.na(groupped$Type), 'no_type', groupped$Type)

p1 <- groupped %>% filter(period=='1985-1992')%>% 
  ggplot(aes(x='',y=troops, fill=Type))+
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values=group.colors)+
  labs(x='', y='', title='1985-1992')+
  theme_bw()+
  theme(axis.text = element_blank(),
        axis.title.x =element_blank(),
        axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
        axis.ticks = element_blank()
       )



p2 <- groupped %>% filter(period=='1993-2000')%>% 
  ggplot(aes(x='',y=troops, fill=Type))+
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values=group.colors)+
  labs(x='', y='', title='1993-2000')+
  theme_bw()+
  theme(axis.text = element_blank(),
        axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
        axis.ticks = element_blank()
        )

p3  <- groupped %>% filter(period=='2001-2013')%>% 
  ggplot(aes(x='',y=troops, fill=Type))+
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values=group.colors)+
  labs(x='', y='', title='2001-2013')+
  theme_bw()+
  theme(axis.text = element_blank(),
        axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
        axis.ticks = element_blank())


p4  <- groupped %>% filter(period=='2014-2022')%>% 
  ggplot(aes(x='',y=troops, fill=Type))+
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values=group.colors)+
  labs(x='', y='', title='2014-2022')+
  theme_bw()+
  theme(axis.text = element_blank(),
        axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
        axis.ticks = element_blank())


ol1 <- ggarrange(p1,p2, p3, p4,
                    ncol = 2, nrow = 2)


ol1


ggsave("~/Articles/images/Figure4.jpg", units = "cm",dpi=600)
ggsave("~/Articles/images/Figure4.eps", units = "cm",dpi=600)

```






### Figure 5 

*Military deployment down by the mandate of organizations and states in four periods*


```{r, message=FALSE,warning=FALSE}
operations %>% 
  pivot_longer(cols=c('UN':'Other'),
                names_to='Mandate',
               values_to = 'freq')->longed_org
```


```{r, message=FALSE,warning=FALSE}

group.colors <- c(UN = "#4363d8",  Other="gray", 
       no_mandate='#303030',           
  US ='red',         
  EU ='#ffe119'  ,       
  NATO ="#004586"  ,     
  SADC ="#bfef45" ,      
  AU    ="#3cb44b" ,     
  ECOWAS ="#9A6324"     ,
  OSCE  ="#f032e6"      ,
 ECCAS ="#aaffc3"     , 
 ECOMOG="#800000")

longed_org$Mandate = ifelse(longed_org$freq<1 , 'no_mandate', longed_org$Mandate)

longed_org %>% 
  group_by(period,  Mandate) %>% summarise(troops=mean(troops,na.rm=T)) %>% 
  arrange(desc(period)) ->groupped

groupped$Mandate=ifelse(is.na(groupped$Mandate), 'no_mandate', groupped$Mandate)

p1 <- groupped %>% filter(period=='1985-1992')%>% 
  ggplot(aes(x='',y=troops, fill=Mandate))+
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values=group.colors)+
  labs(x='', y='', title='1985-1992')+
  theme_bw()+
  theme(axis.text = element_blank(),
        axis.title.x =element_blank(),
        axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
        axis.ticks = element_blank()
       )



p2 <- groupped %>% filter(period=='1993-2000')%>% 
  ggplot(aes(x='',y=troops, fill=Mandate))+
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values=group.colors)+
  labs(x='', y='', title='1993-2000')+
  theme_bw()+
  theme(axis.text = element_blank(),
        axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
        axis.ticks = element_blank()
        )

p3  <- groupped %>% filter(period=='2001-2013')%>% 
  ggplot(aes(x='',y=troops, fill=Mandate))+
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values=group.colors)+
  labs(x='', y='', title='2001-2013')+
  theme_bw()+
  theme(axis.text = element_blank(),
        axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
        axis.ticks = element_blank())


p4  <- groupped %>% filter(period=='2014-2022')%>% 
  ggplot(aes(x='',y=troops, fill=Mandate))+
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values=group.colors)+
  labs(x='', y='', title='2014-2022')+
  theme_bw()+
  theme(axis.text = element_blank(),
        axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
        axis.ticks = element_blank())

ol2 <- ggarrange(p1,p2, p3, p4,
                    ncol = 2, nrow = 2)


ol2


ggsave("~/Articles/images/Figure5.jpg",dpi=800)
ggsave("~/Articles/images/Figure5.eps",dpi=800)

```








### Figure 6

*Figure 7. The military deployment network’s graph metrics in 1985-2022*

 - Density -  the ratio between the number of edges present in the graph and the maximum possible number of edges that the graph can have.
 
 - Transitivity - 
 
 - Reciprocity -
 
 - Modularity - 



```{r, message=FALSE,warning=FALSE}
library(ggraph)
library(tidygraph)
library(igraph)
years <- c(1985:2022)


density_vector <- c()

for(x in years){
  balance %>% filter(year==x) %>% 
    dplyr::select(cowc1,cowc2,troops) %>% na.omit() %>% 
    graph_from_data_frame() %>% graph.density()->new_value
  density_vector <- c(density_vector,new_value)
  
}


transitivity_vector <- c()

for(i in years){
  balance %>% filter(year==i) %>% 
    dplyr::select(cowc1,cowc2,troops) %>% na.omit() %>% 
    graph_from_data_frame() %>% transitivity()->new_value
  transitivity_vector <- c(transitivity_vector,new_value)
  
}

reciprocity_vector <- c()

for(i in years){
  balance %>% filter(year==i) %>% 
    dplyr::select(cowc1,cowc2,troops) %>% na.omit() %>% 
    graph_from_data_frame() %>% reciprocity()->new_value
  reciprocity_vector <- c(reciprocity_vector,new_value)
  
}

modularity_vector <- c()

for(i in years){
  balance %>% filter(year==i) %>% 
    dplyr::select(cowc1,cowc2,troops) %>% na.omit() %>% 
    graph_from_data_frame() %>% as.undirected() %>% cluster_louvain() %>% modularity()->new_value
  modularity_vector <- c(modularity_vector,new_value)
  
}

mod_table <- data.frame(years,density_vector,reciprocity_vector,modularity_vector,
                        transitivity_vector)

mod_table %>% 
  pivot_longer(
    cols =c('density_vector':'transitivity_vector'),
    names_to = 'graph metric',
    values_to = 'value'
  )->graph_metric_long


graph_metric_long %>% 
  ggplot(aes(years, value, color=`graph metric`))+
  geom_line()+
  labs(y='Graph metrics',x='Year', title='Dynamics of graph metrics')+
  scale_fill_manual(values= colors3)+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))+
  theme_bw()+
   theme(axis.title.x = element_text(size=8, face='bold'),axis.title.y = element_text(size=8, face='bold'), title =element_text(size=10, face='bold') )+
  theme(axis.text.x = element_text( size=6))+
  theme(axis.text.y = element_text( size=6))
  

ggsave('~/Articles/images/Figure6.jpg', dpi=800)
ggsave('~/Articles/images/Figure6.eps', dpi=800)

```

